\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-04-20 trunk/arch.tex
updates

In such a process, one challenge is to avoid being overwhelmed by
goals that may threaten the ability of a reactor to process and
integrate in its evolving partial plan. For that reason, each reactor
specifies the following two parameters that applied towards the scope
of the agent lifetime:

@@

The choice of representation of state variables, observations and
goals is directly derived from \eu timelines and tokens.  While such a
rich representation allows exchange of state information in a formal
yet flexible manner, its translation into planning frameworks that
have an explicit representation of time (such as but not limited to
\eu) is often trivial.

@@

Both these functions have different scope and temporal constraints.
The \texttt{synchronize} call is considered as atomic and will focus
only on state inference for the current tick by identifying the
current value of {\em internal} state variables from the latest
updates on {\em external} state variables. This process propagates up
in the reactor dependency graph at every tick to ensure that all
reactors are aware of the current state of the world. \texttt{step}
relates to a single decision step of the planning process focused on
the future evolution of state variables; it can take multiple steps
spanning multiple ticks for a reactor to produce a plan that is both
complete and valid. How the reactor identifies that its plan is
complete and valid is its responsibility and depends on the way each
reactor resolves its plan at each {\em step}. The general idea is that
the reactor is able to find a valid strategy in order to satisfy as
many of its {\em internal} goals as possible. For \eu, this means that
the solver is unable to identify any further flaws to be resolved in
the plan.

@@

The critical task of a \rx reactor is to successfully identify its
state through synchronization as the execution frontier advances. When
a reactor fails to integrate its {\em external} state update and/or
identify its {\em internal} state at any tick, it implies that it is
unable to explain the state of the world.  This in turn implies that
any state information from this reactor is suspect, resulting in the
exclusion of this reactor from the execution cycle and consequent
destruction. As the agent destroys the faulty reactor, a special
observation, \texttt{Failed}, is produced on all {\em internal}
timelines. Reactors which declare these very state variables as {\em
  external}, are informed of the failure of this timeline.  All goals
exchanged between reactors are considered \emph{rejectable} by the
reactor that manages them. As a result in such a situation, the
failure to find a plan for these goals, is \emph{not} considered a
critical failure of the reactor. It will only result on these goals
being ignored until newer goals are received.

@@

This suggests a deliberate non overlap between {\em reactive planner}
and {\em deliberative planners} horizons. In \rxe, and more
specifically in our \eu reactor the planning horizon of {\em
  synchronization} -- limited to the current tick -- is a subset of
the planning horizon of the deliberation steps -- starting at the
current tick and with an extent based on latency $\lambda$ and
look-ahead $\pi$ of the reactor. In this manner we avoid creating a
conflict between the deliberation and synchronization by controlling
the scheduling of these two process by the \rx agent.

@@

If \texttt{object.location} is {\em external} and \texttt{Agent} is an
{\em internal} timeline, a robot may encounter the situation where it
is at the location where it can buy one (possibly among others)
product. This results in a decision-making problem related to the
state value of the \texttt{Agent} timeline; the robot can either {\em
  Buy} a specific product now, wait a little longer before doing so or
just simply not buy any product at this location.

@@

More generally tying a token to future outcomes in a model that is
meant to be confronted by execution is problematic. It might be valid
if one can guarantee that an outcome is predisposed to occur (for
example the reactor managing the \texttt{Agent.location} timeline
guarantees that the outcome of the \texttt{Going} will {\em always}
result on ending at the traget location). However, given the very
nature of the planning problem and the ability to search outcomes and
to replan, the model should be written so that a future projection can
be manipulated.

@@

As the example above shows, synchronization involves a deliberation
process. The risk is such that the search for a solution is not a
simple sequence of decisions where the resolution can be easily
identified by the solver as any other option would lead immediately to
n plan inconsistency. Since such a process is required by the \rx at
\emph{every} tick for \emph{every} reactor the model needs to be
designed carefully.

One critical aspect to take into account is that from a \rx point of
view, synchronization is the only critical task of a reactor. Such a
process is used in order for the agent to identify its current state
and maintain it consistently throughout the agent's lifetime. For a
reactor not being able to find a consistent solution during
synchronization -- implying that after evaluating all possible
solutions provided by a model it was unable to find a consistent plan
-- not only indicates that a reactor is unable to explain the state of
the world, but more importantly it may jeopardize other reactors that
rely on its {\em internal} state. For this reason a failure to
synchronize immediately results in the \rx agent killing a reactor and
notifying any reactor that depends on its {\em internal} state
variables; it does so by posting the special observation
\textsf{Failed}. This special state is a special predicate which is
acceptable for all the {\em internal} and {\em external} timelines of
a \eu reactor.  By so doing we allow for graceful degradation where
other reactors can readapt their plans on receipt of this observation.

@@

\item The last observation has proved that the current plan is invalid
  requiring the {\em synchronization} solver to relax the existing
  plan to identify its current {\em external} state. (This case is
  developed further on the next section).

@@

When the {\em execution frontier} advances, the reactor planning scope
moves further into the future. This is reflected by the {\em
  deliberation} solver which uses a new deliberation scope one tick
further in the future. As Fig. \ref{fig:plan:window} illustrates, such
a scope allows filtering of tokens that are either necessarily ending
before $\tau$ or necessarily starting after, focusing the planning
problem on only those tokens that are within a ``reasonable'' future.
This filtering window complements the \rx {\em dispatch window} from
Definition \ref{def:dispatch} which filters out at the agent level
those goals that are before the reactor execution latency or too far
into the future.  The planning window refines this filtering further,
as tokens can be excluded dynamically from deliberation as their
insertion in the plan and scheduling with other generated tokens can
alter their temporal scope to be outside of the planning window.  This
allows the planner to focus on its planning horizon and report
decision on flaws that are beyond this scope. As, time advances, these
potential future flaws will eventually be part of the planning window
of the reactor and will be then resolved along with potentially new
goals.  This in turn, results in an apparent \emph{continuous
  planning} of the reactor. When there are no more flaws present in
the current planning scope a partial plan solution is found. This has
two effects:

>>>>

In such a process, one challenge is to avoid being overwhelmed by
goals that may threaten the ability of a reactor to process and
integrate in its evolving partial plan. For that reason, each reactor
specifies the following two parameters that \kcomment{are} applied
towards the scope of the agent lifetime:

@@

The choice of representation of state variables, observations and
goals is directly derived from \eu timelines and tokens.  While such a
rich representation allows exchange of state information in a formal
yet flexible manner, its translation into planning frameworks that
have an explicit representation of time (such as but not limited to
\eue) is often trivial.

@@

Both these functions have different scope and temporal constraints.
The \texttt{synchronize} call is considered as atomic and will focus
only on state inference for the current tick by identifying the
current value of {\em internal} state variables from the latest
updates on {\em external} state variables. This process propagates up
in the reactor dependency graph at every tick to ensure that all
reactors are aware of the current state of the world. \texttt{step}
relates to a single decision step of the planning process focused on
the future evolution of state variables; it can take multiple steps
spanning multiple ticks for a reactor to produce a plan that is both
complete and valid. How the reactor identifies that its plan is
complete and valid is its responsibility and depends on the way each
reactor resolves its plan at each {\em step}. The general idea is that
the reactor is able to find a valid strategy in order to satisfy as
many of its {\em internal} goals as possible. For \eue, this means
that the solver is unable to identify any further flaws to be resolved
in the plan.

@@

The critical task of a \rx reactor is to successfully identify its
state through synchronization as the execution frontier advances. When
a reactor fails to integrate its {\em external} state update and/or
identify its {\em internal} state at any tick, it implies that it is
unable to explain the state of the world.  This in turn implies that
any state information from this reactor is suspect, resulting in the
exclusion of this reactor from the execution cycle and consequent
destruction. As the agent destroys the faulty reactor, a special
observation, \texttt{Failed}, is produced on all {\em internal}
timelines. Reactors which declare these very state variables as {\em
  external}, are informed of the failure of this timeline.  All goals
exchanged between reactors are considered \emph{rejectable} by the
reactor that manages them. As a result in such a situation, the
failure to find a plan for these goals, is \emph{not} considered a
critical failure of the reactor. It will only result \kcomment{in}
these goals being ignored until newer goals are received.

@@

This suggests a deliberate non overlap between \kcomment{the} {\em
  reactive planner} and {\em deliberative planners} horizons. In \rxe,
and more specifically in our \eu reactor the planning horizon of {\em
  synchronization} -- limited to the current tick -- is a subset of
the planning horizon of the deliberation steps -- starting at the
current tick and with an extent based on latency $\lambda$ and
look-ahead $\pi$ of the reactor. In this manner we avoid creating a
conflict between the deliberation and synchronization by controlling
the scheduling of these two process\kcomment{es} by the \rx agent.

@@

If \texttt{object.location\kcomment{.At}} is {\em external} and \texttt{Agent} is an
{\em internal} timeline, a robot may encounter the situation where it
is at the location where it can buy one (possibly among others)
product. This results in a decision-making problem related to the
state value of the \texttt{Agent} timeline; the robot can either {\em
  Buy} a specific product now, wait a little longer before doing so or
just simply not buy any product at this location.

@@

More generally tying a token to future outcomes in a model that is
meant to be confronted by execution is problematic. It might be valid
if one can guarantee that an outcome is predisposed to occur (for
example the reactor managing the \texttt{Agent.location} timeline
guarantees that the outcome of the \texttt{Going} will {\em always}
result on ending at the \kcomment{target} location). However, given
the very nature of the planning problem and the ability to search
outcomes and to replan, the model should be written so that a future
projection can be manipulated.

@@

As the example above shows, synchronization involves a deliberation
process. The risk is such that the search for a solution is not a
simple sequence of decisions where the resolution can be easily
identified by the solver as any other option would lead immediately to
plan inconsistency. Since such a process is required by the \rx at
\emph{every} tick for \emph{every} reactor the model needs to be
designed carefully.

One critical aspect to take into account is that from a \rx point of
view, synchronization is the only critical task of a reactor. Such a
process is used in order for the agent to identify its current state
and maintain it consistently throughout the agent's lifetime. For a
reactor not being able to find a consistent solution during
synchronization -- implying that after evaluating all possible
solutions provided by a model it was unable to find a consistent plan
-- not only indicates that a reactor is unable to explain the state of
the world, but more importantly it may jeopardize other reactors that
rely on its {\em internal} state. For this reason a failure to
synchronize immediately results in the \rx agent killing a reactor and
notifying any reactor that depends on its {\em internal} state
variables; it does so by posting the special observation
\textsf{Failed}. This special state is a special predicate which is
acceptable for all the {\em internal} and {\em external} timelines of
\kcomment{an} \eu reactor.  By so doing we allow for graceful
degradation where other reactors can readapt their plans on receipt of
this observation.

@@

\item The last observation has proved that the current plan is invalid
  requiring the {\em synchronization} solver to relax the existing
  plan to identify its current {\em external} state. (This case is
  developed further \kcomment{in} the next section).

@@

When the {\em execution frontier} advances, the reactor planning scope
moves further into the future. This is reflected by the {\em
  deliberation} solver which uses a new deliberation scope one tick
further in the future. As Fig. \ref{fig:plan:window} illustrates, such
a scope allows filtering of tokens that are either necessarily ending
before $\tau$ or necessarily starting after, focusing the planning
problem on only those tokens that are within a ``reasonable'' future.
This filtering window complements the \rx {\em dispatch window} from
Definition \ref{def:dispatch} which filters out at the agent level
those goals that are before the reactor execution latency or too far
into the future.  The planning window refines this filtering further,
as tokens can be excluded dynamically from deliberation as their
insertion in the plan and scheduling with other generated tokens can
alter their temporal scope to be outside of the planning window.  This
allows the planner to focus on its planning horizon and report
decision on flaws that are beyond this scope. As time advances, these
potential future flaws will eventually be part of the planning window
of the reactor and will be then resolved along with potentially new
goals.  This in turn, results in an apparent \emph{continuous
  planning} of the reactor. When there are no more flaws present in
the current planning scope a partial plan solution is found. This has
two effects:
